---
title: "ADVTTEST R package"
author: "Caxton Muchono"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{advttest}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(ADVTTEST)
```


# Theory

The use of t.test theory implemented in this package. Specifically, the package 
will test the NULL hypothesis that the underlying means of two populations are
the same. That is you will be testing:

$H_0 : \mu_x = \mu_y$

or more usefully

$H_1 :  \mu_x - \mu_y = 0$

This involves using a built-in function called `t.test(x,y)` where x and y are 
samples from the respective populations. Some important experimental design 
structures must be taken into consideration before using the function. These
considerations will be expressed through options supplied to the t.test function.

### Considerations

If x and y are independent and the variances in the populations are equal 
then the test is performed using `var.equal = TRUE`

`t.test(x,y, var.equal = TRUE)`

If x and y are independent and the population variances are not equal 
then the test is done using 

`t.test(x,y, var.equal = FALSE)`

If the samples are "paired", that is each experimental unit is measured 
twice, the first goes into x and the last into y then the appropriate test is

`t.test(x,y,paired=TRUE)`

The NULL hypothesis is placed in the options as $\mu=0$ which is the 
default value.


##  The myttest function 

Myttest function will investigate the data for evidence for a difference
in the population variances if the data supports the conclusion that the 
underlying variances are likely different or if paired = TRUE then an 
appropriate t-test is made to produce a list containing:

- "Welch", "T-test", "Paired" depending on what test was conducted
- A "Y/N" conclusion as to whether we should reject the NULL of equality 
of means at the alpha level of significance
- The summary statistics made by the appropriate t.test in a data list

The function will takes two vectors, alpha and paired. Assume that the 
variances in the population are the same (this will mean a simple no 
conditional function)

The function must produce a named list containing:

* A data frame containing x and y
* alpha $\alpha$
* paired
* The confidence interval for $\mu_x - \mu_y$
* The pvalue
* the test type
* the "Y/N" conclusion

The list will be of class "Rttest"

```{r, eval=FALSE}

myttest = function(x, y, alpha, paired){
  t.test = NULL

  # create a data frame
  df = data.frame(x = x, y = y)

  # finding the population variance
  sd1 = sd(x)
  sd2 = sd(y)

  # determine conclusion that the underlying variances
  if(paired)                   { test_type = "Paired"; v = FALSE }
  else if (abs(sd1 - sd2) < 5) { test_type = "T-test"; v = TRUE  }
  else                         { test_type = "WELCH "; v = FALSE }

  # make t-test assuming that the variances in the population are the same
  tests = t.test(x = x, y = y,
                 var.equal = v, paired = paired, conf.level = 1 - alpha)

  # find the confidence intervals and p-value
  conf_inter = tests$conf.int
  p_value = tests$p.value

  if(p_value < alpha) { conclusion = "Y" }
  else                { conclusion = "N" }

  # make a named list of function
  fcn_list = list(data = df, alpha = alpha, paired = paired,
                  conf_inter = conf_inter, p_value = p_value,
                  test_type = test_type, conclusion = conclusion)

  # make list to be of class "Rttest"
  class(fcn_list) = "Rttest"

  # return the list
  return(fcn_list)
}

```

## An informative plot with the "Rttest" class

The "Rttest" class that creates boxplots of the data when 
non - paired data is analyzed. It also will make a boxplot of differences in 
data when `paired = TRUE` The ci for the difference of means is plotted inside
the boxplot of differences.

```{r, eval=FALSE}

plot.Rttest = function(x, ...){

  if(x$paired){   # if paired

    data_var = NULL    # the letter of the the data (X)
    value    = NULL    # value of data

    # merge the differences in data of x object
    m_data = data.frame(value = x$data$x - x$data$y, data_var = rep(LETTERS[24]))

    # create the boxplot
    ggplot(m_data, aes(x = data_var , y= value , fill=data_var )) +
      geom_boxplot() +
      stat_summary(fun=mean, geom="point", shape=23, size=4, fill="white") +
      scale_fill_discrete(name = "categories") +
      labs(x = "Difference", y = "samples diff",
           title ="Boxplot of the samples")
  }
  else{           # else if not paired

    data_var = NULL    # the letter of the the data (X or Y)
    value    = NULL    # value of data

    # create separate data sets of x and y
    x_data = data.frame(value = x$data$x, data_var = rep(LETTERS[24]))
    y_data = data.frame(value = x$data$y, data_var = rep(LETTERS[25]))

    # merge the x and y data into new data set
    merge_data = merge(x = x_data, y = y_data, all = TRUE)

    # create the boxplot
    ggplot(merge_data, aes(x = data_var , y= value , fill=data_var )) +
      scale_fill_discrete(name = "categories") +
      geom_boxplot() +
      labs(x = "category", y = "samples",
           title ="Boxplot of the samples")
  }
}

```

## An informative print with the "Rttest" class

Print function prints to the command line  A (1-$\alpha$)100% confidence 
interval for $\mu_x - \mu_y$ and the type of test

```{r, eval=FALSE}

print.Rttest = function(x, ...) {

  ci        = x$conf_inter  # store the confidence interval
  test_type = x$test_type   # store the type of test

  # create a list with summary data and return it
  pr_list = list(ci = ci, test_type = test_type)

  return(pr_list)
}

```

### Test the constructor with a function call from the provided data

```{r}

set.seed(32); x=rnorm(30,mean=10,sd=15)
set.seed(35); y=rnorm(30,mean=8,sd=15)
ans1=ADVTTEST::myttest(x,y,alpha=0.05,paired=FALSE)
print(ans1)
plot(ans1)

set.seed(32); x=rnorm(30,mean=10,sd=5)
set.seed(35); y=rnorm(30,mean=8,sd=15)
ans2=ADVTTEST::myttest(x,y,alpha=0.05,paired=FALSE)
print(ans2)
plot(ans2)

set.seed(32); x=rnorm(30,mean=10,sd=15)
set.seed(35); y = x+ rnorm(30, 5 ,4)
ans3=ADVTTEST::myttest(x,y,alpha=0.05,paired=TRUE)
print(ans3)
plot(ans3)

```
